---
title: "R: Basic Operations"
output: html_notebook
---

# R: Basic Operations

## Comments

Comments in R are denoted by `#`

```{r}
# This is a comment
```

There is no built-in way to do multi-line comments.

## Mathematical Operations

Mathematical operations in R are handled using the standard 4 symbols `+, -, /, and *` . Operations are calculated using PEMDAS.

### Addition

```{r}
# Addition
1 + 2
```

### Subtraction

```{r}
# Subtraction
3 - 1
```

### Multiplication

```{r}
# Multiplication
10 * 42
```

### Division

```{r}
# Division
120 / 12
```

## Variable Assignment

Variables in R are traditionally assigned using an arrow sign `<-` , made using a less than carat and a dash. It is valid to use `=` , but not recommended.

```{r}
x <- 12
x
```

Variable names can not include and spaces or symbols other than the underscore `_` . They cannot begin with numbers, but can include numbers after the first character.

```{r}
variable_name1 = 5
variable_name1
```

## Data Types

R has the following data types: numeric, integer, character, logical, and complex.

### Numeric

Decimal values in R are numerics. Numerics are the default data type for numbers.

```{r}
a <- 10
a
```

```{r}
b <- 1.5
b
```

Note: Numerics are not integers. You can confirm this with both a and b above using the `is.integer()` function.

```{r}
class(a)
```

```{r}
class(b)
```

### Integer

To create an integer in R we can use the `as.integer()` function.

```{r}
c <- as.integer(1)
is.integer(c)
```

```{r}

```

For shorthand, we can also define an integer by writing the desired integer value, followed by a capital L.

```{r}
d <- 15L
```

```{r}
class(d)
```

Integers must be whole numbers, if passed a floating value, the number will be rounded to an integer.

```{r}
c <- as.integer(3.7)
c
```

This also works with a string.

```{r}

```

We would use an integer data-type when we want to represent a whole integer, and don't want to accept any other variations of numbers. The numeric data-type is less strict, it allows us to switch between different number types like integers and floats dynamically.

### Character

A **character** type is used to represent strings in R. Characters can be denoted by writing within quotes.

```{r}
# Defining a character object
name <- "Joe"
name
```

#### Concatenating character objects

Character objects can be concatenated into a single variable using the `paste` function

```{r}
first_name = "Joe"
last_name = "Smith"
full_name = paste(first_name, last_name)
full_name
```

#### Converting into character objects

We can convert other objects into character objects using the `as.character()` function

```{r}
x <- as.character(12)
x
```

```{r}
# Note that class returns the data type of our object
class(x)
```

#### Printing character types

You can print out hardcoded character objects with the `print` function

```{r}
print("John ate 12 cookies")
```

You can print out formatted character objects with the `sprintf` function. You can use the following symbols to represent formatted parts of the code, each should be preceded by a `%` symbol:

-   d and i - Integer values

-   s - Character string

-   o - Octal

-   x and X - Hexadecimal

-   f - Double precision value (For fixed point representation, for example: 21.234)

-   e and E - Double precision value (For exponential representation, for example: 2.1234e+01)

```{r}
sprintf("%s ate %d cookies.", "John", 21)
```

```{r}

```

#### Return a substring

You can get a substring from a character object using the `substr` function

```{r}
substr("How much wood could a woodchuck chuck?", start <- 5, stop <- 13)
```

#### Replacing characters in a string

\+

You can replace characters in a string using the `sub` function

```{r}
sub("much", "little", "How much wood could a woodchuck chuck?")
```

### Logical

A logical value is commonly referred to as a boolean value. It is the result of a comparison between two variables. It is represented as `TRUE` or `FALSE.`

```{r}
12 < 10
```

```{r}
10 <= 12
```

#### Logical Operators

The standard logical operations are:

-   `&` - And

-   `|` - Or

-   `!` - Not

```{r}
x <- TRUE
y <- FALSE
```

```{r}
x & y
```

```{r}
x | y
```

```{r}
!x
```

### Complex

A complex object in R is created using the symbol `i` . The use of the i symbol is in reference to imaginary numbers.

```{r}

```

```{r}

```

## Comparison Operators

You can use the following operators to compare two variables and return a logical object:

-   `<` Less than

-   `>` Greater than

-   `<=` Less than or equal to

-   `>=` Greater than or equal to

-   `==` Is equal to

-   `!=` Is NOT equal to

```{r}
10 < 12
```

```{r}
10 > 12
```

```{r}
10 <= 12
```

```{r}
10 >= 12
```

```{r}
10 == 12
```

```{r}
10 != 12
```

## Homework

1.  Create two strings with your first and last name, respectively. Then concatenate them, and print the new value as "My name is \<your newly created concatenated variable will go here\>"

    ```{r}
    first_name <- "Nick"
    last_name <- "Hammer"
    full_name <- paste(first_name, last_name)
    sprintf("My name is %s", full_name)

    ```

2.  What values can a BOOLEAN contain?

```{r}
TRUE
FALSE

```
